Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-04-03 document/ms.tex
correct formulation of GP prior

% Document specific variables.
\newcommand{\rate}{\ensuremath{\eta}}
\newcommand{\ratepars}{{\ensuremath{\bvec{\theta}}}}
\newcommand{\obs}[1]{\ensuremath{\hat{#1}}}
\newcommand{\radius}{\ensuremath{R}}
\newcommand{\period}{\ensuremath{T}}
\newcommand{\completeness}{{\ensuremath{P_\mathrm{c}}}}
\newcommand{\transitprob}{{\ensuremath{P_\mathrm{t}}}}
\newcommand{\entry}{\ensuremath{S}}
\newcommand{\catalog}{{\ensuremath{\bvec{\entry}}}}
\newcommand{\binheight}{{\ensuremath{w}}}
\newcommand{\binheights}{{\ensuremath{\bvec{\binheight}}}}

\newcommand{\smoothreg}{{\ensuremath{\lambda}}}
\newcommand{\interim}{{\ensuremath{\bvec{\alpha}}}}

@@

In the following sections, we will model the occurrence rate
$\rate(\period,\,\radius)$ as a two dimensional grid with $N_\period$ bins in
$\ln\period$ and $N_\radius$ bins in $\ln\radius$.
We choose to fix the bin locations leaving only the bin heights
$\binheights = \{\binheight_{ij}\}_{i=1,j=1}^{N_\period,N_\radius}$ as
free parameters.

@@

Since the choice of bin locations and number is arbitrary, the most
conservative strategy is to choose very small bins---yielding an effectively
non-parametric model---and choose a prior (or regularization) that encodes any
prior expectations about the distribution.
In particular, it seems reasonable to expect the distribution to be
``smooth''.
The most extreme version of this model would be a completely non-parametric
rate function---generated by, for example, a Gaussian process
(\citealt{poiss-gp})---but this discussion is beyond the scope of this
\paper.
Instead, we'll simply choose small bins and add a Gaussian process prior
(\dfmtodo{CITE}) on the logarithms of the bin heights
\begin{eqnarray}
\ln p(\binheights\,|\,) &=&
-\frac{1}{2}\,\bvec{r}^\mathrm{T}\,\bvec{K}\,\bvec{r}
-\frac{1}{2}\ln \det \bvec{K} - \frac{N_\period\,N_\radius}{2}\ln 2\pi
\end{eqnarray}
where $\bvec{r}$ is the vector of residuals
\begin{eqnarray}
\bvec{r} &=& \left(\begin{array}{cccc}
\binheight_1-\mu & \binheight_2-\mu & \cdots & \binheight_N-\mu
\end{array}\right)^\mathrm{T}
\end{eqnarray}
and $\bvec{K}$ is the kernel matrix generated by
\begin{eqnarray}
K_{ij} &=&
\end{eqnarray}

>>>>

% Document specific variables.
\newcommand{\rate}{\ensuremath{\eta}}
\newcommand{\ratepars}{{\ensuremath{\bvec{\theta}}}}
\newcommand{\obs}[1]{\ensuremath{\hat{#1}}}
\newcommand{\radius}{\ensuremath{R}}
\newcommand{\period}{\ensuremath{T}}
\newcommand{\completeness}{{\ensuremath{P_\mathrm{c}}}}
\newcommand{\transitprob}{{\ensuremath{P_\mathrm{t}}}}
\newcommand{\entry}{\ensuremath{S}}
\newcommand{\catalog}{{\ensuremath{\bvec{\entry}}}}

\newcommand{\interim}{{\ensuremath{\bvec{\alpha}}}}

\newcommand{\bincenter}{{\ensuremath{\bvec{x}}}}
\newcommand{\binheight}{{\ensuremath{w}}}
\newcommand{\binheights}{{\ensuremath{\bvec{\binheight}}}}

\newcommand{\smooth}{{\ensuremath{\lambda}}}

@@

In the following sections, we will model the occurrence rate
$\rate(\period,\,\radius)$ as a two dimensional grid with $N$ bins in
$\ln\period$ and $\ln\radius$.
We choose to fix the bin locations leaving only the bin heights
$\binheights = \{\binheight_i\}_{i=1}^{N}$ as free parameters.

@@

Since the choice of bin locations and number is arbitrary, the most
conservative strategy is to choose very small bins---yielding an effectively
non-parametric model---and choose a prior (or regularization) that encodes any
prior expectations about the distribution.
In particular, it seems reasonable to expect the distribution to be
``smooth''.
The most extreme version of this model would be a completely non-parametric
rate function---generated by, for example, a Gaussian process
(\citealt{poiss-gp})---but this discussion is beyond the scope of this
\paper.
Instead, we'll simply choose small bins and add a Gaussian process prior
(\dfmtodo{CITE}) on the logarithms of the bin heights
\begin{eqnarray}
\ln p(\binheights\,|\,\mu,\,\bvec{\smooth}) &=&
-\frac{1}{2}\,\bvec{r}^\mathrm{T}\,\bvec{K}^{-1}\,\bvec{r}
-\frac{1}{2}\ln \det \bvec{K} - \frac{N}{2}\ln 2\pi
\end{eqnarray}
where $\bvec{r}$ is the vector of residuals
\begin{eqnarray}
\bvec{r} &=& \left(\begin{array}{cccc}
\binheight_1-\mu & \binheight_2-\mu & \cdots & \binheight_N-\mu
\end{array}\right)^\mathrm{T}
\end{eqnarray}
and $\bvec{K}$ is the kernel matrix evaluated \emph{at the bin centers}
$\bincenter_i$
\begin{eqnarray}
K_{ij}(\bvec{\smooth}) &\equiv& \smooth_0^2\,\exp \left(
-\frac{1}{2}[\bincenter_i-\bincenter_j]^\mathrm{T}
\left[\begin{array}{cc}
1/\smooth_1 & 0 \\ 0 & 1/\smooth_2
\end{array}\right]
[\bincenter_i-\bincenter_j] \right) \quad.
\end{eqnarray}

Below, we will explore the effects of this choice of prior but when we
generate our final results, we will actually apply hyperpriors to the
parameters $\mu$ and $\bvec{\smooth}$ and marginalize over their values.
